---
title: "DATA_605_Assignment_10_DThonn"
file: DThonn_Assign10_DATA_605
name: Daniel Thonn
output:
  html_document: default
  pdf_document: default
course: DATA 605 CUNY MSDA program
---

** DATA_605_Assignment_10_Thonn - Markov Chains  **


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
# install libraries if needed
#install.packages("permutations")
#library(permutations)

#install.packages('gtools')
#library(gtools)

```
Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if
(a) he bets 1 dollar each time (timid strategy).
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
(c) Which strategy gives Smith the better chance of getting out of jail? 

** Assignment 10 - Ch.11 **
(a) Timid: he bets 1 dollar each time (timid strategy).

This situation is a Markov chain that has 9 steps (0 to 8)

See page 418 of the text (Intro to Probability, Grinstead,Snell) describing the Fundamental Matrix
for a Markov Absorbing Chain

$$ if~in~the~1~state~ s_j ~after ~k~steps:~ P(X^k = 1)= q_{ij}^k$$
$$ if~in~the~0~state~ s_j ~after ~k~steps: 1 - q_{ij}^k$$

state_0 = $0
state_1 = $1
state_2 = $2
state_3 = $3
state_4 = $4
state_5 = $5
state_6 = $6
state_7 = $7
state_8 = $8




``````{r, echo=TRUE}

PWin <- 0.4
PLoss <- 1 - PWin
state1 <- 1
state8 <- 8

pr_timid <- (1 - (PLoss/PWin)^state1)/(1 - (PLoss/PWin)^state8)
pr_timid

#Result: [1] 0.02030135


```

(b) Bold: he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).

For maximun win the Markav Chain must be: s=1 to s=2 to s=4 to arrive at s=8.

``````{r, echo=TRUE}

PWin <- 0.4
PLoss <- 1 - PWin
state1b <- 1
state2b <- 2

#for each step of 1 to the next step 2

ps1 <- (1 - (PLoss/PWin)^state1b)/(1 - (PLoss/PWin)^state2b)
ps1

ps2 <- (1 - (PLoss/PWin)^state1b)/(1 - (PLoss/PWin)^state2b)
ps2

ps4 <- (1 - (PLoss/PWin)^state1b)/(1 - (PLoss/PWin)^state2b)
ps4

Ps1_ps8 <-ps1 * ps2 * ps4 
Ps1_ps8
# Result: [1] 0.064


```

(c) Which strategy gives Smith the better chance of getting out of jail? 

Examining the results of (a) and (b) it can be seen it is best to go wit the Bold strategy with higher probability.

** END **


``````{r, echo=TRUE}

```


